---
title: "PhysioTwin: the movement digital-twin pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PhysioTwin: the movement digital-twin pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

A *digital twin* couples a mechanistic model (whose parameters are physical or
clinical quantities) with realistic sensor emulation and data assimilation, so a
subject-specific model can be **personalised** from real measurements and then
used for **in-silico experiments**. This vignette runs the four layers
end-to-end on the movement modality --- the reference vertical slice --- using
only synthetic data, so it is fully reproducible offline.

```{r setup}
library(PhysioTwin)
set.seed(1)
```

## 1--2. A "real" subject, recorded through a noisy IMU

Layer 1 is a mechanistic single-joint limb whose parameters map to clinical
quantities (muscle `strength`, passive `damping`, joint `stiffness`). Layer 2
turns the clean simulated motion into a realistic inertial-measurement-unit
stream with bias and noise.

```{r simulate}
subject <- limbTwin(strength = 5.5, damping = 0.7, stiffness = 3.2)
truth   <- simulateTwin(subject, duration = 4, dt = 0.02, theta0 = 0.1)
rec     <- imuMeasure(truth,
                      imuSensor(gyro_bias = 0.015, gyro_noise = 0.035,
                                accel_noise = 0.12))
head(rec, 3)
```

## 3. Personalise the twin from the recording

Layer 3 assimilates the IMU stream into a twin, starting from a *wrong* prior
and recovering the subject's parameters. `n_starts` is reduced here to keep the
vignette fast; the default (25) is more robust.

```{r personalize}
prior <- limbTwin(strength = 4, damping = 0.3, stiffness = 3.2)
twin  <- personalizeTwin(rec, prior, estimate = c("strength", "damping"),
                         dt = 0.02, n_starts = 3)
twin$twin[c("strength", "damping")]   # compare with the true 5.5 and 0.7
```

## 4a. In-silico intervention

With a personalised twin, perturb parameters and read off the predicted change:
here a rehabilitation that adds 40% strength and removes 30% damping.

```{r intervention}
effect <- insilicoIntervention(twin$twin,
                               scale = list(strength = 1.4, damping = 0.7),
                               duration = 5, dt = 0.01)
effect$outcomes
```

## 4b. A labelled synthetic training set

The same generators produce domain-randomised, labelled data for machine
learning --- each row is one simulated subject with its true parameters and
signal features.

```{r trainingdata}
train <- generateTrainingData(n = 24, duration = 4, dt = 0.02, seed = 7)
dim(train)
head(as.data.frame(train)[, c("strength", "damping", "stiffness")], 3)
```

## Validity and verification

Simulation validity is not the same as matching one dataset. `validateTwin()`
separates **verification** (are the equations solved correctly?) from
**validation** (are they the right equations?) and reports the operating
envelope. The cheap checks are shown here; `recovery`, `sensitivity` and
`prediction` (which refit the twin many times) are available via the `checks`
argument.

```{r validate}
validateTwin(limbTwin(strength = 6, damping = 0.5, stiffness = 6),
             checks = c("verification", "identifiability", "domain"))
```

## Where next

The physiological generators (`jansenRit()` for EEG, `ecgsyn()` /
`windkessel()` for the cardiovascular system, `respiration()`, `kuramoto()`)
and the wider estimator family (`unscentedKalmanFilter()`, `particleFilter()`,
`abcSMC()`, `gpCalibrate()`) are the building blocks for coupling the other
modalities into this same loop. See `?PhysioTwin` for the full map.
